---
title: "Case Study 1:  Exploratory Data Analysis (EDA) for Budweiser"
author: "Phu Truong & Kristi Herman"
date: "01/16/2020"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

### The purpose of this exploratory data analysis is to answer questions from Budweiser related to the craft beer industry in the United States.

```{r}
# Import dependencies
library(tidyverse)
library(dplyr)
library(ggplot2)
library(data.table)
library(ggthemes)
library(usmap)
library(naniar)
library(class)
library(caret)

# Import Beer and Brewery data
beer_df <-  read_csv("Beers.csv")
brew_df <-  read_csv("Breweries.csv")

# Check to make sure data imported
dim(beer_df)
dim(brew_df)
View(brew_df)

```

## 1.  How many breweries are present in each state?

```{r}
#  Group breweries by state
brew_by_state <- brew_df %>% group_by(State) %>% summarize(count=n())

# Display # of breweries by state
View(brew_by_state)

# Plot # of breweries by state in a barchart
brew_by_state %>%
  ggplot(aes(x = State, y = count)) +
  geom_bar(stat = "identity", fill = "blue", alpha = .5) +
  ggtitle("Number of Breweries by State") + xlab("State") + ylab("Number of Breweries") +
  geom_text(aes(y = count, label = count), fontface = "bold", vjust = -.4, color = "black", size = 3) 

# Need to fix this plot of breweries by state in a map
# plot_usmap(data = brew_by_state$State, values = brew_by_state$count, color = "red") + 
#   scale_fill_continuous(name = "Population (2015)", label = scales::comma) + 
#   theme(legend.position = "right")

```

## 2.  Merge the Beer Data with the Brewery Data.  Print first and last six observations.

```{r}
# Change the column names in the Beer df
setnames(beer_df, old=c("Name","Brewery_id"), new=c("Beer_name", "Brew_ID"))

# Change column name in the Brewery df
setnames(brew_df, old=c("Name"), new=c("Brewery_name"))

# Make the state column a factor
brew_df$State <-  factor(brew_df$State)
str(brew_df)

# Merge the brewery data with the beer data (left join)
all_df <- merge(beer_df, brew_df, by="Brew_ID")
dim(all_df)

# Print the first & last 6 observations
(first6= head(all_df,6))
(last6= tail(all_df,6))
```

## 3.  Address the missing values in each column

```{r}

# count missing values in each column
s = sapply(all_df, function(x) sum(is.na(x)))
s

gg_miss_var(all_df)

# Remove missing data for beer & rename the Name and Brewery_id Columns
clean_df <- all_df %>% filter(!is.na(ABV) & !is.na(IBU) & !is.na(Style))
dim(clean_df)
```

## 4.  Compute the median alcohol content and international bitterness unit for each state.  Plot a bar chart to compare.

```{r}
# Dataframe with median values for ABV and IBU
(med_abv_ibu <- clean_df %>% 
  group_by(State) %>%
  summarize(med_abv = median(ABV), med_ibu = median(IBU), count = n()))

# Plot
ggplot(med_abv_ibu, aes(x = State)) +
  geom_col(aes( y = med_ibu, fill="redfill")) +
  geom_text(aes(y = med_ibu, label = med_ibu), fontface = "bold", vjust = 2.4, color = "white", size = 3) +
  geom_line(aes(y = med_abv * 1000, group = 1, color = 'blackline')) +
  geom_text(aes(y = med_abv * 1000, label = round(med_abv, 2)), vjust = -.4, color = "black", size = 3) +
  scale_y_continuous(sec.axis = sec_axis(trans = ~ . / 500)) +
  scale_fill_manual('', labels = 'Median Bitterness (IBU)', values = "#C00000") +
  scale_color_manual('', labels = 'Median Alcohol By Volumn (ABV)', values = 'black') +
  ggtitle("Median Alcohol Content & International Bitterness by State") +
  theme_minimal()

```

## 5.  Which state has the maxiumum alcoholic (ABV) beer?  Which state has the most bitter (IBU) beer?
###Kentucky has the highest ABV beer, which is called ‘London Balling’,  made by Against the Grain Brewery, located in Louisville city.
###Oregon has the highest IBU beer, which is called ‘Bitter Bitch Imperial IPA’, made by Astoria Brewing Company, located in Astoria city.
```{r}
max_abv_ibu <- clean_df %>% group_by(State) %>% summarize(max_abv = max(ABV), max_ibu = max(IBU), count = n())

# State with max ABV
(highest_abv <- max_abv_ibu %>% filter(rank(desc(max_abv))<=1))

# State with max IBU
(highest_ibu <- max_abv_ibu %>% filter(rank(desc(max_ibu))<=1))

#List row with max ABV
view(clean_df[which.max(clean_df$ABV),])

#List row with max IBU
view(clean_df[which.max(clean_df$IBU),])



```

## 6.  Comment on the summary statistics and distribution of the ABV variable.

### The distribution of ABV variable is slightly right-skewed.

```{r}
#Summary of key statistics of ABV
summary(clean_df$ABV)

# Plot boxplot
clean_df %>%
  ggplot(aes(y = ABV)) +
  geom_boxplot() +
  ggtitle("Summary Statistics of the ABV") + ylab("ABV")

# Plot histogram
clean_df %>%
  ggplot(aes(x = ABV)) +
  geom_histogram(colour="black",fill="navy") +
  ggtitle("Distribution of ABV") + xlab("ABV")

```

## 7.  Is there an apparant relationship between the bitterness of the beer and its alcoholic content?  Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.

### With p-value < 2.2e-16, there is sufficient evidence at alpha= .05 level of significance to suggest that the data is linearly correlated. Correlation estimate =0.67 suggests that the relationship between IBU and ABV is positive and strong.

```{r}
# Plot ABV and IBU
ggplot(data = clean_df) +
  geom_point(mapping = aes(x = ABV, y = IBU, color=ABV), position ="jitter") +
  ggtitle("Alcohol Content & Bitterness of Beers") + xlab("Alcohol by Volume (ABV)") + ylab("International Bitterness Units (IBU)")

#correlation
cor.test(clean_df$ABV,clean_df$IBU)

```

## 8.  Investigate the difference in IBU and ABV between IPAs and other Ales

```{r}

# Create a dataframe with only IPAs & Ales
ipa_ale_df <- clean_df %>% filter(str_detect(Style, 'IPA') |str_detect(Style, 'Ale'))

# Add a column for "IPA" or "Ale"
ipa_ale_df$beer_type = ifelse(grepl("IPA", ipa_ale_df$Style), "IPA", "Ale")

# Check to make sure there are no NAs for IPA or Ale
ipa_ale_df %>% filter(is.na(beer_type))
str(ipa_ale_df)

# Split data into train and test data
# Train/Test with 70/30
splitPerc = .7
set.seed(4)

trainIndices = sample(1:dim(ipa_ale_df)[1],round(splitPerc * dim(ipa_ale_df)[1]))
train = ipa_ale_df[trainIndices,]
test = ipa_ale_df[-trainIndices,]

dim(train)
head(train)
dim(test)
head(test)
```

```{r}
# Run the KNN model on the train and test data with k = 5
classifications = knn(train[,c("ABV","IBU")],test[,c("ABV","IBU")],train$beer_type, prob = TRUE, k = 5)

# Create a confusion matrix of the results with k = 5
table(classifications,test$beer_type)
CM_KNN_IPA = confusionMatrix(table(classifications,test$beer_type))
CM_KNN_IPA

# Predict what Budweiser would be classified as
bud_test = data.frame(ABV = .05, IBU = 12)

classify_bud = knn(train[,c("ABV","IBU")], bud_test, train$beer_type, prob = TRUE, k = 5)
classify_bud

#IPA vs Ale
# Loop for many k and the average of many training / test partition
# to pick the best k for the model 
```{r}
iterations = 200
numks = 30

masterAcc = matrix(nrow = iterations, ncol = numks)
  
for(j in 1:iterations)
{
accs = data.frame(accuracy = numeric(30), k = numeric(30))
trainIndices = sample(1:dim(ipa_ale_df)[1],round(splitPerc * dim(ipa_ale_df)[1]))
train = ipa_ale_df[trainIndices,]
test = ipa_ale_df[-trainIndices,]
for(i in 1:numks)
{
   classifications = knn(train[,c("ABV","IBU")],test[,c("ABV","IBU")],train$beer_type, prob = TRUE, k = i)
  table(test$beer_type,classifications)
  CM = confusionMatrix(table(test$beer_type,classifications))
  masterAcc[j,i] = CM$overall[1]
}

}

MeanAcc = colMeans(masterAcc)

plot(seq(1,numks,1),MeanAcc, type = "l")

# Plot scatter plot of ABV and IBU for beer classified as “Ale” and “IPA”
ggplot(data = ipa_ale_df) +
  geom_point(mapping = aes(x = ABV, y = IBU, color=beer_type), position ="jitter") +
  ggtitle("IPA and Ales") + xlab("Alcohol by Volume (ABV)") + ylab("International Bitterness Units (IBU)")

```

## 9.  What states are closest to the Bud ABV/IBU profile?

```{r}
# Run the KNN model on the train and test data and add the state parameter (Use all data to train/test)
classify_w_state = knn(train[, 4:5],test[, 4:5],train$State, prob = TRUE, k = 5)
classify_bud_st = knn(train[, 4:5], bud_test, train$State, prob = TRUE, k = 5)

CM_KNN_State = confusionMatrix(table(classify_w_state,test$State))
CM_KNN_State

```